\documentclass{Configuration_Files/PoliMi3i_thesis}

%------------------------------------------------------------------------------
%	REQUIRED PACKAGES AND  CONFIGURATIONS
%------------------------------------------------------------------------------

% CONFIGURATIONS
\usepackage{parskip} % For paragraph layout
\usepackage{setspace} % For using single or double spacing
\usepackage{emptypage} % To insert empty pages
\usepackage{multicol} % To write in multiple columns (executive summary)
\usepackage{numprint} % Print formatted numbers
\setlength\columnsep{15pt} % Column separation in executive summary
\setlength\parindent{0pt} % Indentation
\usepackage{subfiles} % Include sub files
\usepackage{comment} % Comments
\raggedbottom  

% PACKAGES FOR TITLES
\usepackage{titlesec}
% \titlespacing{\section}{left spacing}{before spacing}{after spacing}
\titlespacing{\section}{0pt}{3.3ex}{2ex}
\titlespacing{\subsection}{0pt}{3.3ex}{1.65ex}
\titlespacing{\subsubsection}{0pt}{3.3ex}{1ex}
\usepackage{color}

% PACKAGES FOR LANGUAGE AND FONT
\usepackage[english]{babel} % The document is in English  
\usepackage[utf8]{inputenc} % UTF8 encoding
\usepackage[T1]{fontenc} % Font encoding
\usepackage[11pt]{moresize} % Big fonts

% PACKAGES FOR IMAGES
\usepackage{graphicx}
\usepackage{transparent} % Enables transparent images
\usepackage{eso-pic} % For the background picture on the title page
\usepackage{subfig} % Numbered and caption subfigures using \subfloat.
\usepackage{tikz} % A package for high-quality hand-made figures.
\usetikzlibrary{}
\graphicspath{{./Images/}} % Directory of the images
\usepackage{caption} % Coloured captions
\usepackage{xcolor} % Coloured captions
\usepackage{amsthm,thmtools,xcolor} % Coloured "Theorem"
\usepackage{float}

% STANDARD MATH PACKAGES
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{bm}
\usepackage[overload]{empheq} % For braced-style systems of equations.
\usepackage{fix-cm} % To override original LaTeX restrictions on sizes

% PACKAGES FOR TABLES
\usepackage{tabularx}
\usepackage{longtable} % Tables that can span several pages
\usepackage{colortbl}

% PACKAGES FOR ALGORITHMS (PSEUDO-CODE)
\usepackage{algorithm}
\usepackage{algorithmic}

% PACKAGES FOR REFERENCES & BIBLIOGRAPHY
\usepackage[colorlinks=true,linkcolor=black,anchorcolor=black,citecolor=black,filecolor=black,menucolor=black,runcolor=black,urlcolor=black]{hyperref} % Adds clickable links at references
\usepackage{cleveref}
\usepackage[square, numbers, sort&compress]{natbib} % Square brackets, citing references with numbers, citations sorted by appearance in the text and compressed
\bibliographystyle{abbrvnat} % You may use a different style adapted to your field

% OTHER PACKAGES
\usepackage{pdfpages} % To include a pdf file
\usepackage{afterpage}
\usepackage{lipsum} % DUMMY PACKAGE
\usepackage{fancyhdr} % For the headers
\fancyhf{}

\usepackage{listings} % code snippets
\usepackage{courier}
\lstset{framextopmargin=50pt,frame=bottomline}
\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{mymauve}{rgb}{0.58,0,0.82}
\lstset{ %
  backgroundcolor=\color{white},   % choose the background color
  basicstyle=\footnotesize\ttfamily,        % size of fonts used for the code
  breaklines=true,                 % automatic line breaking only at whitespace
  captionpos=b,                    % sets the caption-position to bottom
  commentstyle=\color{mygreen},    % comment style
  escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
  keywordstyle=\color{blue},       % keyword style
  stringstyle=\color{mymauve},     % string literal style
  frame=none                       % no lower frame
}

% Input of configuration file. Do not change config.tex file unless you really know what you are doing. 
\input{Configuration_Files/config}

%----------------------------------------------------------------------------
%	NEW COMMANDS DEFINED
%----------------------------------------------------------------------------

% EXAMPLES OF NEW COMMANDS
\newcommand{\bea}{\begin{eqnarray}} % Shortcut for equation arrays
\newcommand{\eea}{\end{eqnarray}}
\newcommand{\e}[1]{\times 10^{#1}}  % Powers of 10 notation

%----------------------------------------------------------------------------
%	ADD YOUR PACKAGES (be careful of package interaction)
%----------------------------------------------------------------------------

%----------------------------------------------------------------------------
%	ADD YOUR DEFINITIONS AND COMMANDS (be careful of existing commands)
%----------------------------------------------------------------------------

%----------------------------------------------------------------------------
%	BEGIN OF YOUR DOCUMENT
%----------------------------------------------------------------------------

\begin{document}

\fancypagestyle{plain}{%
\fancyhf{} % Clear all header and footer fields
\fancyhead[RO,RE]{\thepage} %RO=right odd, RE=right even
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}}

%----------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------

\pagestyle{empty} % No page numbers
\frontmatter % Use roman page numbering style (i, ii, iii, iv...) for the preamble pages

\puttitle{
	title=Titolo da scegliere, % Title of the thesis
	name=Ciriello Giovanni, % Author Name and Surname
	course=ICT Engineering\, Business and Innovation - Computer Science Engineering, % Study Programme (in Italian)
	ID  = 963100,  % Student ID number (numero di matricola)
	advisor= Prof. Francalanci Chiara, % Supervisor name
	coadvisor={Ravanelli Paolo}, % Co-Supervisor name, remove this line if there is none
	academicyear={2021-22},  % Academic Year
} % These info will be put into your Title page 

%----------------------------------------------------------------------------
%	PREAMBLE PAGES: ABSTRACT (inglese e italiano), EXECUTIVE SUMMARY
%----------------------------------------------------------------------------
\startpreamble
\setcounter{page}{1} % Set page counter to 1

% ABSTRACT IN ENGLISH
\chapter*{Abstract} 
Maritime transport has always been considered the most effective and widespread means of transferring substantial cargoes of goods in terms of size and weight over long distances. Since it constitutes 90\% of the world's trade \cite{trasporto-marittimo}, security is one of the most discussed aspects, and recently an intense research activity has concerned this field in recent times. 
In this context, more and more agencies are deploying control algorithms that exploit artificial intelligence and machine learning techniques in order to monitor maritime voyages. ARCOS (Arctic Observatory for Copernicus SEA) is a European Union project that aims, among other things, to develop predictive warning systems that can ensure continuous monitoring of the Arctic region. For this purpose, it has collected data about the routes taken by ships in the area of concern. These data contain messages that comply with the Automatic Identification System (AIS) standard, which is crucial for ensuring the safety of maritime operations. The system works with numerous sensors on board and a transceiver that sends the data - at a frequency of the order of minutes - to orbiting satellites (S-AIS messages). The scope of this thesis is to propose a methodology to detect anomalies of any kind from a dataset of AIS messages. In order to achieve this goal, the messages were first grouped into trips, that represent the entity on which anomalies were chosen to be detected. A spatial density-based clustering algorithm (DBSCAN) was applied to them, followed by techniques to analyze the resulting outliers to assess their relevance and significance. The method was validated on a test dataset, containing more than 95 million messages, where potential anomalies were detected in terms of distance traveled, navigation time, and load variations. A possible future development is the creation of a predictive warning system that - taking advantage of the proposed methodology - would allow satellites to detect anomalies as messages are received as a preliminary step to a manual check by the relevant authorities.
\\
\\
\textbf{Keywords:} AIS, Anomaly, Maritime Transport, DBSCAN, Monitoring.

% ABSTRACT IN ITALIAN
\chapter*{Abstract in lingua italiana}

Il trasporto marittimo è storicamente considerato il mezzo più efficace e diffuso per trasferire su lunga distanza consistenti carichi di merci in termini di dimensione e peso. Costituendo la base del 90\% del commercio mondiale \cite{trasporto-marittimo}, la sicurezza rappresenta uno degli aspetti più discussi e una intensa attività di ricerca ha interessato questo campo negli ultimi tempi.  
In tale contesto, sempre più enti stanno implementando algoritmi di controllo che sfruttano tecniche di intelligenza artificiale e machine learning al fine di monitorare i viaggi marittimi. ARCOS (Arctic Observatory for Copernicus SEA) è un progetto finanziato dall'Unione Europea che ha, tra i suoi obiettivi, quello di sviluppare sistemi di allerta predittiva in grado di fornire un continuo monitoraggio della regione Artica. A tale scopo, ha collezionato dati relativi ai tragitti percorsi nella zona interessata da mezzi di trasporto marittimo. Tali dati contengono messaggi che rispettano lo standard AIS (Automatic Identification System), cruciale per garantire la sicurezza delle operazioni marittime. Il suo funzionamento si basa su numerosi sensori presenti a bordo e una ricetrasmittente che invia i dati - con una frequenza dell’ordine di minuti - a satelliti in orbita (messaggi S-AIS). Lo scopo della tesi è quello di proporre una metodologia per rilevare anomalie di qualsiasi genere a partire da un dataset di messaggi AIS. Al fine di raggiungere tale obiettivo, i messaggi sono stati in primo luogo raggruppati in tragitti, che rappresentano l'entità sulla quale si è scelto di rilevare anomalie. Su di essi è stato applicato un algoritmo di clustering basato su densità spaziale (DBSCAN) e successivamente tecniche di analisi dei valori anomali risultanti al fine di valutarne la rilevanza e il significato. Il metodo è stato validato su un dataset di testing, contenente più di 95 milioni di messaggi, sul quale sono state rilevate potenziali anomalie relative a distanza percorsa, tempi di navigazione e variazioni del carico. Un possibile sviluppo futuro è la creazione di un sistema di allerta predittiva che - sfruttando la metodologia proposta - permetta ai satelliti il rilevamento di anomalie in fase di ricezione dei messaggi come step preliminare ad un controllo manuale da parte delle autorità competenti.

\\
\\
\textbf{Parole chiave:} AIS, Anomalia, Trasporto Marittimo, DBSCAN, Monitoraggio.

%----------------------------------------------------------------------------
%	LIST OF CONTENTS/FIGURES/TABLES/SYMBOLS
%----------------------------------------------------------------------------

% TABLE OF CONTENTS
\thispagestyle{empty}
\tableofcontents % Table of contents 
\thispagestyle{empty}
\cleardoublepage

%-------------------------------------------------------------------------
%	THESIS MAIN TEXT
%-------------------------------------------------------------------------
% In the main text of your thesis you can write the chapters in two different ways:
%
%(1) As presented in this template you can write:
%    \chapter{Title of the chapter}
%    *body of the chapter*
%
%(2) You can write your chapter in a separated .tex file and then include it in the main file with the following command:
%    \chapter{Title of the chapter}
%    \input{chapter_file.tex}
%
% Especially for long thesis, we recommend you the second option.

\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics
\mainmatter % Begin numeric (1,2,3...) page numbering

% --------------------------------------------------------------------------
% NUMBERED CHAPTERS % Regular chapters following
% --------------------------------------------------------------------------

\input{chapters/0_introduction}

\input{chapters/1_stateOfTheArt}
    

\input{chapters/2_methodology}


\input{chapters/3_testing}


\input{chapters/4_conclusions}
    
%-------------------------------------------------------------------------
%	BIBLIOGRAPHY
%-------------------------------------------------------------------------

\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics
\bibliography{Thesis_bibliography} % The references information are stored in the file named "Thesis_bibliography.bib"

%-------------------------------------------------------------------------
%	APPENDICES
%-------------------------------------------------------------------------

\cleardoublepage
\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics
\appendix

\input{appendices/A_data-description}

\input{appendices/B_testing-results}

\input{appendices/C_pca}


% LIST OF FIGURES
\listoffigures

% LIST OF TABLES
\listoftables

% ACKNOWLEDGEMENTS

\chapter*{Acknowledgements}



\cleardoublepage

\end{document}
